---
title: 'Global surface temperature change detection: Differences between the long-term mean of 1950-1970 and 1990-2010'
author:
- "Merlin Unterfinger, Geography UZH"
- "Olivier Niklaus, Geography UZH"
date: 4 7 2017
output:
  rmarkdown::pdf_document:
    fig_caption: yes        
    includes:  
      in_header: preamble-latex.tex
    number_sections: yes
    toc: yes
    toc_depth: 3
  html_document: default
abstract: "Spatial continuity of global temperature point measurements is analysed by creating H-scatterplots, autocovariance plots and variograms. Universal Kriging is applied to interpolate global temperature maps from the point measurements. Global layers of elevation, sun inclination angle, atmospheric distance and continentality are created and used in the Universal Kriging as explaining variables. Afterwards difference images are calculated and interpreted."
filter: pandoc-eqnos
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())             # Clean the environment
options(scipen=6)         # Display digits, not the scientific version
options(digits.secs=6)    # use milliseconds in Date/Time data types
options(warning=FALSE)    # Don't show warnings
par(mfrow=c(1,1))         # reset plot placement to normal 1 by 1

dataFolder    <- file.path(getwd(), "data")            # Data folder
resultsFolder <- file.path(getwd(), "results")         # Result folder
figureFolder  <- file.path(getwd(), "figures")         # Figure Folder
```

\newpage
```{r Libraries, message=FALSE, echo=FALSE}
library(ape)
library(dichromat)
library(FNN)
library(ggplot2) 
library(gdistance)
library(gridExtra)
library(gstat)
library(raster)
library(rasterVis)
library(rgdal)
library(rgeos)
library(sp)
```

# Data
```{r Geodata, message=FALSE, warning=FALSE , echo=FALSE, fig.align="center"}
# Define CRS
WGS84 <- CRS("+init=epsg:4326")

# Read CSV file for temperature
temp <- read.csv(file.path(dataFolder, "temperature.csv"))
temp <- temp[complete.cases(temp), ]

# Train and validation data
selection <- sample(1:nrow(temp), floor(nrow(temp)*0.05), replace = FALSE)
temp.val    <- temp[selection,]
temp  <- temp[-selection,]

# Create spdf's and assign CRS
coordinates(temp.val)<-~long+lat
proj4string(temp.val)<-WGS84
coordinates(temp)<-~long+lat
proj4string(temp)<-WGS84

# Read test file
temp.test <- read.csv(file.path(dataFolder, "temperature_test.csv"))
coordinates(temp.test) <- ~long+lat
proj4string(temp.test) <- WGS84

plot(temp, main="All measurements and validation points")
points(temp.val, col="red")

knitr::kable(head(round(as.data.frame(head(temp)),2), caption = "Structure of the temperature data set.", row.names = FALSE))

# Seperate data according to timeFrame and season
temp1970w <-SpatialPointsDataFrame(coords = temp@coords, data = temp@data[,1:2])
temp1970s <-SpatialPointsDataFrame(coords = temp@coords, data = temp@data[,c(1,3)])
temp2010w <-SpatialPointsDataFrame(coords = temp@coords, data = temp@data[,c(1,4)])
temp2010s <-SpatialPointsDataFrame(coords = temp@coords, data = temp@data[,c(1,5)])

col <- c("id", "meansum")
colnames(temp1970s@data) <- col
colnames(temp2010s@data) <- col
colnames(temp1970w@data) <- col
colnames(temp2010w@data) <- col

# Open digital elevation model
DEM <-readGDAL(file.path(dataFolder, "globalDHM.tif"), silent = T)
#plot(DEM, col=topo.colors(20))
landDEM <- raster(DEM)
landDEM[landDEM<0] <- 0

# Set up grid, for prediction
grid <- spsample(DEM,type="regular",100)
grid <- SpatialPointsDataFrame(grid@coords,data.frame(elevation=extract(landDEM,grid)))

# Load coastlines for plotting
oceans <- readOGR(dsn="data", "ne_110m_ocean", verbose=F)

# Plot
p <- levelplot(raster(DEM),
          main = "DEM",
          par.settings=rasterTheme(region = dichromat(topo.colors(7))), margin=F,
          at=seq(-maxValue(abs(raster(DEM))), maxValue(abs(raster(DEM))), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p)

# Save plots
pdf(file.path(figureFolder, "DEM.pdf"), width=10.5, height=7.5)
print(p)
invisible(dev.off())

# Clear plots
remove(p)
```

\newpage
# Spatial continuity
## H-scatterplots and autocovariance
### Winter before 1970
```{r scatter1970w, echo=FALSE, fig.align="center", fig.height=4}
temp.dist<-spDists(temp1970w,longlat = FALSE)
temp.index <-knn.index(temp.dist, k=20, algorithm=c("kd_tree"))

temp.h<-data.frame(sum=temp1970w$meansum ,sumNN1=temp1970w$meansum[temp.index[,1]],sumNN5=temp1970w$meansum[temp.index[,5]],sumNN10=temp1970w$meansum[temp.index[,10]],sumNN20=temp1970w$meansum[temp.index[,20]])

par(mfrow=c(2,2))
plot(temp.h$sum,temp.h$sumNN1,main="h = 1")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN5,main="h = 5")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN10,main="h = 10")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN20,main="h = 20")
abline(1,1, col = "blue")

# Autocoavariance plots
temp.mean<-mean(temp.h$sum, na.rm = TRUE)
temp.sd<-sd(temp.h$sum, na.rm = TRUE)

temp.acov.nn1<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN1-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn5<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN5-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn10<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN10-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn20<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN20-temp.mean), na.rm = TRUE)/nrow(temp.h)

temp.acorr.nn1<-temp.acov.nn1/temp.sd^2
temp.acorr.nn5<-temp.acov.nn5/temp.sd^2
temp.acorr.nn10<-temp.acov.nn10/temp.sd^2
temp.acorr.nn20<-temp.acov.nn20/temp.sd^2

par(mfrow=c(1,2))
plot(c(1,5,10,20),c(temp.acov.nn1,temp.acov.nn5,temp.acov.nn10,temp.acov.nn20),type="l",main="auto-covariance")
plot(c(1,5,10,20),c(temp.acorr.nn1,temp.acorr.nn5,temp.acorr.nn10,temp.acorr.nn20),type="l",main="auto-correlation")

```

### Winter after 1990
```{r scatter2010w, echo=FALSE, fig.align="center", fig.height=4}
temp.dist<-spDists(temp2010w,longlat = FALSE)
temp.index <-knn.index(temp.dist, k=20, algorithm=c("kd_tree"))

temp.h<-data.frame(sum=temp2010w$meansum ,sumNN1=temp2010w$meansum[temp.index[,1]],sumNN5=temp2010w$meansum[temp.index[,5]],sumNN10=temp2010w$meansum[temp.index[,10]],sumNN20=temp2010w$meansum[temp.index[,20]])

par(mfrow=c(2,2))
plot(temp.h$sum,temp.h$sumNN1,main="h = 1")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN5,main="h = 5")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN10,main="h = 10")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN20,main="h = 20")
abline(1,1, col = "blue")

# Autocoavariance plots
temp.mean<-mean(temp.h$sum, na.rm = TRUE)
temp.sd<-sd(temp.h$sum, na.rm = TRUE)

temp.acov.nn1<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN1-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn5<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN5-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn10<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN10-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn20<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN20-temp.mean), na.rm = TRUE)/nrow(temp.h)

temp.acorr.nn1<-temp.acov.nn1/temp.sd^2
temp.acorr.nn5<-temp.acov.nn5/temp.sd^2
temp.acorr.nn10<-temp.acov.nn10/temp.sd^2
temp.acorr.nn20<-temp.acov.nn20/temp.sd^2

par(mfrow=c(1,2))
plot(c(1,5,10,20),c(temp.acov.nn1,temp.acov.nn5,temp.acov.nn10,temp.acov.nn20),type="l",main="auto-covariance")
plot(c(1,5,10,20),c(temp.acorr.nn1,temp.acorr.nn5,temp.acorr.nn10,temp.acorr.nn20),type="l",main="auto-correlation")
```

### Summer before 1970
```{r scatter1970s, echo=FALSE, fig.align="center", fig.height=4}

temp.dist<-spDists(temp1970s,longlat = FALSE)
temp.index <-knn.index(temp.dist, k=20, algorithm=c("kd_tree"))

temp.h<-data.frame(sum=temp1970s$meansum ,sumNN1=temp1970s$meansum[temp.index[,1]],sumNN5=temp1970s$meansum[temp.index[,5]],sumNN10=temp1970s$meansum[temp.index[,10]],sumNN20=temp1970s$meansum[temp.index[,20]])

par(mfrow=c(2,2))
plot(temp.h$sum,temp.h$sumNN1,main="h = 1")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN5,main="h = 5")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN10,main="h = 10")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN20,main="h = 20")
abline(1,1, col = "blue")

# Autocovariance plots
temp.mean <- mean(temp.h$sum, na.rm = TRUE)
temp.sd<-sd(temp.h$sum, na.rm = TRUE)

temp.acov.nn1<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN1-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn5<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN5-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn10<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN10-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn20<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN20-temp.mean), na.rm = TRUE)/nrow(temp.h)

temp.acorr.nn1<-temp.acov.nn1/temp.sd^2
temp.acorr.nn5<-temp.acov.nn5/temp.sd^2
temp.acorr.nn10<-temp.acov.nn10/temp.sd^2
temp.acorr.nn20<-temp.acov.nn20/temp.sd^2

par(mfrow=c(1,2))
plot(c(1,5,10,20),c(temp.acov.nn1,temp.acov.nn5,temp.acov.nn10,temp.acov.nn20),type="l",main="auto-covariance")
plot(c(1,5,10,20),c(temp.acorr.nn1,temp.acorr.nn5,temp.acorr.nn10,temp.acorr.nn20),type="l",main="auto-correlation")

```

### Summer after 1990
```{r scatter2010s, echo=FALSE, fig.align="center", fig.height=4}

temp.dist<-spDists(temp2010s,longlat = FALSE)
temp.index <-knn.index(temp.dist, k=20, algorithm=c("kd_tree"))

temp.h<-data.frame(sum=temp2010s$meansum ,sumNN1=temp2010s$meansum[temp.index[,1]],sumNN5=temp2010s$meansum[temp.index[,5]],sumNN10=temp2010s$meansum[temp.index[,10]],sumNN20=temp2010s$meansum[temp.index[,20]])

par(mfrow=c(2,2))
plot(temp.h$sum,temp.h$sumNN1,main="h = 1")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN5,main="h = 5")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN10,main="h = 10")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN20,main="h = 20")
abline(1,1, col = "blue")

# Autocoavariance plots
temp.mean<-mean(temp.h$sum, na.rm = TRUE)
temp.sd<-sd(temp.h$sum, na.rm = TRUE)

temp.acov.nn1<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN1-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn5<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN5-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn10<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN10-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn20<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN20-temp.mean), na.rm = TRUE)/nrow(temp.h)

temp.acorr.nn1<-temp.acov.nn1/temp.sd^2
temp.acorr.nn5<-temp.acov.nn5/temp.sd^2
temp.acorr.nn10<-temp.acov.nn10/temp.sd^2
temp.acorr.nn20<-temp.acov.nn20/temp.sd^2

par(mfrow=c(1,2))
plot(c(1,5,10,20),c(temp.acov.nn1,temp.acov.nn5,temp.acov.nn10,temp.acov.nn20),type="l",main="auto-covariance")
plot(c(1,5,10,20),c(temp.acorr.nn1,temp.acorr.nn5,temp.acorr.nn10,temp.acorr.nn20),type="l",main="auto-correlation")

```

## Empirical Variogram
### Winter before 1970
```{r temp1970wvar, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp1970w.var1<-variogram(meansum~1,temp1970w,cutoff=1000,width=1)
temp1970w.var5<-variogram(meansum~1,temp1970w,cutoff=150,width=5)
temp1970w.var10<-variogram(meansum~1,temp1970w,cutoff=150,width=10)
temp1970w.var15<-variogram(meansum~1,temp1970w,cutoff=150,width=15)

grid.arrange(plot(temp1970w.var1, main="Bin width 1km"), 
             plot(temp1970w.var5, main="Bin width 5km"), ncol = 2)

grid.arrange(plot(temp1970w.var10, main="Bin width 10km"), 
             plot(temp1970w.var15, main="Bin width 15km"), ncol = 2)

bws<-data.frame(np=temp1970w.var1$np,bw=rep("1km",length(temp1970w.var1$np)))
bws<-rbind(bws,data.frame(np=temp1970w.var5$np,bw=rep("10km",length(temp1970w.var5$np))))
bws<-rbind(bws,data.frame(np=temp1970w.var10$np,bw=rep("20km",length(temp1970w.var10$np))))
bws<-rbind(bws,data.frame(np=temp1970w.var15$np,bw=rep("30km",length(temp1970w.var15$np))))

#plotting the numbers of points per bin and bandwidth as a boxplot
ggplot(bws, aes(x=bw, y=np, color=bw)) +
  geom_boxplot()
```

\newpage
### Winter after 1990
```{r temp2010wvar, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp2010w.var1<-variogram(meansum~1,temp2010w,cutoff=1000,width=1)
temp2010w.var5<-variogram(meansum~1,temp2010w,cutoff=150,width=5)
temp2010w.var10<-variogram(meansum~1,temp2010w,cutoff=150,width=10)
temp2010w.var15<-variogram(meansum~1,temp2010w,cutoff=150,width=15)

grid.arrange(plot(temp2010w.var1, main="Bin width 1km"), 
             plot(temp2010w.var5, main="Bin width 5km"), ncol = 2)

grid.arrange(plot(temp2010w.var10, main="Bin width 10km"), 
             plot(temp2010w.var15, main="Bin width 15km"), ncol = 2)

bws<-data.frame(np=temp2010w.var1$np,bw=rep("1km",length(temp2010w.var1$np)))
bws<-rbind(bws,data.frame(np=temp2010w.var5$np,bw=rep("10km",length(temp2010w.var5$np))))
bws<-rbind(bws,data.frame(np=temp2010w.var10$np,bw=rep("20km",length(temp2010w.var10$np))))
bws<-rbind(bws,data.frame(np=temp2010w.var15$np,bw=rep("30km",length(temp2010w.var15$np))))

#plotting the numbers of points per bin and bandwidth as a boxplot
ggplot(bws, aes(x=bw, y=np, color=bw)) +
  geom_boxplot()
```

\newpage
### Summer before 1970
```{r temp1970svar, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp1970s.var1<-variogram(meansum~1,temp1970s,cutoff=1000,width=1)
temp1970s.var5<-variogram(meansum~1,temp1970s,cutoff=150,width=5)
temp1970s.var10<-variogram(meansum~1,temp1970s,cutoff=150,width=10)
temp1970s.var15<-variogram(meansum~1,temp1970s,cutoff=150,width=15)

grid.arrange(plot(temp1970s.var1, main="Bin width 1km"), 
             plot(temp1970s.var5, main="Bin width 5km"), ncol = 2)

grid.arrange(plot(temp1970s.var10, main="Bin width 10km"), 
             plot(temp1970s.var15, main="Bin width 15km"), ncol = 2)

bws<-data.frame(np=temp1970s.var1$np,bw=rep("1km",length(temp1970s.var1$np)))
bws<-rbind(bws,data.frame(np=temp1970s.var5$np,bw=rep("10km",length(temp1970s.var5$np))))
bws<-rbind(bws,data.frame(np=temp1970s.var10$np,bw=rep("20km",length(temp1970s.var10$np))))
bws<-rbind(bws,data.frame(np=temp1970s.var15$np,bw=rep("30km",length(temp1970s.var15$np))))

#plotting the numbers of points per bin and bandwidth as a boxplot
ggplot(bws, aes(x=bw, y=np, color=bw)) +
  geom_boxplot()
```

\newpage
### Summer after 1990
```{r temp2010svar, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp2010s.var1<-variogram(meansum~1,temp2010s,cutoff=1000,width=1)
temp2010s.var5<-variogram(meansum~1,temp2010s,cutoff=150,width=5)
temp2010s.var10<-variogram(meansum~1,temp2010s,cutoff=150,width=10)
temp2010s.var15<-variogram(meansum~1,temp2010s,cutoff=150,width=15)

grid.arrange(plot(temp2010s.var1, main="Bin width 1km"), 
             plot(temp2010s.var5, main="Bin width 5km"), ncol = 2)

grid.arrange(plot(temp2010s.var10, main="Bin width 10km"), 
             plot(temp2010s.var15, main="Bin width 15km"), ncol = 2)

bws<-data.frame(np=temp2010s.var1$np,bw=rep("1km",length(temp2010s.var1$np)))
bws<-rbind(bws,data.frame(np=temp2010s.var5$np,bw=rep("10km",length(temp2010s.var5$np))))
bws<-rbind(bws,data.frame(np=temp2010s.var10$np,bw=rep("20km",length(temp2010s.var10$np))))
bws<-rbind(bws,data.frame(np=temp2010s.var15$np,bw=rep("30km",length(temp2010s.var15$np))))

#plotting the numbers of points per bin and bandwidth as a boxplot
ggplot(bws, aes(x=bw, y=np, color=bw)) +
  geom_boxplot()
```

## Fitted Semivariogram
### Winter before 1970
```{r temp1970wfit, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp1970w.var10<-variogram(meansum~1,temp1970w,cutoff=150,width=10)

models <- c("Exp", "Sph", "Gau", "Mat")
temp1970w.var10.fits <- lapply(X=1:4, FUN=function(x) fit.variogram(temp1970w.var10, model=vgm(models[x])))

grid.arrange(plot(temp1970w.var10, model=temp1970w.var10.fits[[1]], main = "Exponential: 10km"), 
             plot(temp1970w.var10, model=temp1970w.var10.fits[[2]], main = "Spherical: 10km"), ncol = 2)
grid.arrange(plot(temp1970w.var10, model=temp1970w.var10.fits[[3]], main = "Gaussian: 10km"), 
             plot(temp1970w.var10, model=temp1970w.var10.fits[[4]], main = "Mat: 10km"), ncol = 2)
```

\newpage
### Winter after 1990
```{r temp2010wfit, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp2010w.var10<-variogram(meansum~1,temp2010w,cutoff=150,width=10)

models <- c("Exp", "Sph", "Gau", "Mat")
temp2010w.var10.fits <- lapply(X=1:4, FUN=function(x) fit.variogram(temp2010w.var10, model=vgm(models[x])))

grid.arrange(plot(temp2010w.var10, model=temp2010w.var10.fits[[1]], main = "Exponential: 10km"), 
             plot(temp2010w.var10, model=temp2010w.var10.fits[[2]], main = "Spherical: 10km"), ncol = 2)
grid.arrange(plot(temp2010w.var10, model=temp2010w.var10.fits[[3]], main = "Gaussian: 10km"), 
             plot(temp2010w.var10, model=temp2010w.var10.fits[[4]], main = "Mat: 10km"), ncol = 2)
```

\newpage
### Summer before 1970
```{r temp1970sfit, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp1970s.var10<-variogram(meansum~1,temp1970s,cutoff=150,width=10)

models <- c("Exp", "Sph", "Gau", "Mat")
temp1970s.var10.fits <- lapply(X=1:4, FUN=function(x) fit.variogram(temp1970s.var10, model=vgm(models[x])))

grid.arrange(plot(temp1970s.var10, model=temp1970s.var10.fits[[1]], main = "Exponential: 10km"), 
             plot(temp1970s.var10, model=temp1970s.var10.fits[[2]], main = "Spherical: 10km"), ncol = 2)
grid.arrange(plot(temp1970s.var10, model=temp1970s.var10.fits[[3]], main = "Gaussian: 10km"), 
             plot(temp1970s.var10, model=temp1970s.var10.fits[[4]], main = "Mat: 10km"), ncol = 2)
```

\newpage
### Summer after 1990
```{r temp2010sfit, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp2010s.var10<-variogram(meansum~1,temp2010s,cutoff=150,width=10)

models <- c("Exp", "Sph", "Gau", "Mat")
temp2010s.var10.fits <- lapply(X=1:4, FUN=function(x) fit.variogram(temp2010s.var10, model=vgm(models[x])))

grid.arrange(plot(temp2010s.var10, model=temp2010s.var10.fits[[1]], main = "Exponential: 10km"), 
             plot(temp2010s.var10, model=temp2010s.var10.fits[[2]], main = "Spherical: 10km"), ncol = 2)
grid.arrange(plot(temp2010s.var10, model=temp2010s.var10.fits[[3]], main = "Gaussian: 10km"), 
             plot(temp2010s.var10, model=temp2010s.var10.fits[[4]], main = "Mat: 10km"), ncol = 2)
```

\newpage
# Universal Kriging
## Continentality
```{r continentality, eval=TRUE, echo=FALSE, warnings=FALSE, message=FALSE, fig.align="center"}
# Load highres coastlines for distance layer
oceans.50m <- readOGR(dsn="data", "ne_50m_ocean", verbose=F)
# Simplify geometry
oceans.simply <-SpatialPolygonsDataFrame(
  gSimplify(oceans.50m, tol=1, topologyPreserve=TRUE),
  data=oceans.50m@data)

# Build blank raster
r <- raster(DEM)
r <- setValues(r, 0)
r <- aggregate(r, fact=10)

# Make values NA where polygon intesects raster
r <- mask(r, oceans.simply)

# Run distance check
ocean.dist <- distance(r)

# # to restrict your search
# searchLimit <- 100 # the maximum distance in raster units from lakes/boundaries
# oceans.simply.buff <- gBuffer(oceans.simply, width = searchLimit, byid = T)
# rB <- crop(r, extent(oceans.simply.buff))
# 
# # much faster...
# rDB <- distance(rB)

levelplot(ocean.dist,
          main = "Continentality",
          par.settings=YlOrRdTheme(), margin=F, 
          at=seq(0, max(abs(cellStats(ocean.dist, range))), len=100)) + 
  layer(sp.polygons(oceans, fill='black', alpha=1))


pdf(file.path(figureFolder, "Continentality.pdf"), width=10.5, height=7.5)
levelplot(ocean.dist,
          main = "continentality",
          par.settings=YlOrRdTheme(), margin=F, 
          at=seq(0, max(abs(cellStats(ocean.dist, range))), len=100)) + 
  layer(sp.polygons(oceans, fill='black', alpha=1))
invisible(dev.off())

```

\newpage
## Surface gradient (North-South)
```{r gradient, eval=TRUE, echo=FALSE,  warnings=FALSE, message=FALSE, fig.align="center"}
# Create latitude and longitude layers
lon <- lat <- raster(DEM)
lon[] <- coordinates(raster(DEM))[, 1]
lat[] <- coordinates(raster(DEM))[, 2]

# Slope and aspect
slope <- terrain(landDEM, opt='slope', unit='radians')
aspect <- terrain(landDEM, opt='aspect', unit='radians')

# Winter
aPer <- (180/2+23)/180
split.hemi.w <- setValues(lat, 1)
split.hemi.w[1:floor((nrow(lat)*aPer)), 1:ncol(lat)] <- -1

# Summer
aPer <- (180/2-23)/180
split.hemi.s <- setValues(lat, 1)
split.hemi.s[1:floor((nrow(lat)*aPer)), 1:ncol(lat)] <- -1

# Calculate the gradient in latitude direction
latGrad <- (cos(aspect)*slope)*180/pi
latGrad.w <- latGrad*split.hemi.w
latGrad.s <- latGrad*split.hemi.s

# Plot
p1 <- levelplot(latGrad.w,
          main = "North-South gradient, +23.5° hemisphere corrected",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(latGrad.w)), maxValue(abs(latGrad.w)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent', alpha=1))
print(p1)

p2 <- levelplot(latGrad.s,
          main = "North-South gradient, -23.5° hemisphere corrected",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(latGrad.s)), maxValue(abs(latGrad.s)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent', alpha=1))
print(p1)

# Save plots
pdf(file.path(figureFolder, "gradient_WiHemiCorr.pdf"), width=10.5, height=7.5)
print(p1)
invisible(dev.off())

pdf(file.path(figureFolder, "gradient_SuHemiCorr.pdf"), width=10.5, height=7.5)
print(p2)
invisible(dev.off())

# Clear plots
remove(p1, p2)
```

## Sun inclination angle
```{r inclination, eval=TRUE, echo=FALSE,  warnings=FALSE, message=FALSE, fig.align="center"}
## Solar inclination angle (height at midday)
# http://www.geoastro.de/astro/mittag/index.htm; h = 90° - β + δ
# Substract ghradient northern hemisphere and add to the southern
# 21.6 - "Summer"; +23.5°
h.w <- (90-abs(lat + 23.5)+latGrad.w)
h.s <- (90-abs(lat - 23.5)+latGrad.s)

# Plot
p1 <- levelplot(h.w,
          main = "Inclination angle sun, winter",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(h.w)), maxValue(abs(h.w)), len=100)) + 
  layer(sp.polygons(oceans, fill='blue', alpha=0.1))
print(p1)

p2 <- levelplot(h.s,
          main = "Inclination angle sun, summer",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(h.w)), maxValue(abs(h.w)), len=100)) + 
  layer(sp.polygons(oceans, fill='blue', alpha=0.1))
print(p2)

# Save plots
pdf(file.path(figureFolder, "inclination_winter.pdf"), width=10.5, height=7.5)
print(p1)
invisible(dev.off())

pdf(file.path(figureFolder, "inclination_summer.pdf"), width=10.5, height=7.5)
print(p2)
invisible(dev.off())
```

\newpage
## Atmospheric distance
```{r d, eval=TRUE, echo=FALSE, warnings=FALSE, message=FALSE, fig.align="center"}
# Create distance layer
lat.w.rad <- abs(lat + 23.5)*pi/180
atmo.dist.w <- 1-cos(lat.w.rad)

lat.s.rad <- abs(lat - 23.5)*pi/180
atmo.dist.s <- 1-cos(lat.s.rad)

# Plot
p1 <- levelplot(atmo.dist.w,
          main = "Atmospheric distance, winter",
          par.settings=BuRdTheme(), margin=T, 
          at=seq(-maxValue(abs(atmo.dist.w)), maxValue(abs(atmo.dist.w)), len=100)) + 
  layer(sp.polygons(oceans, fill='blue', alpha=0.1))
suppressWarnings(print(p1))

p2 <- levelplot(atmo.dist.s,
          main = "Atmospheric distance, summer",
          par.settings=BuRdTheme(), margin=T, 
          at=seq(-maxValue(abs(atmo.dist.s)), maxValue(abs(atmo.dist.s)), len=100)) + 
  layer(sp.polygons(oceans, fill='blue', alpha=0.1))
suppressWarnings(print(p2))

# Save plots
pdf(file.path(figureFolder, "atmoDist_winter.pdf"), width=10.5, height=7.5)
suppressWarnings(print(p1))
invisible(dev.off())

pdf(file.path(figureFolder, "atmoDist_summer.pdf"), width=10.5, height=7.5)
suppressWarnings(print(p2))
invisible(dev.off())

remove(p1, p2)
```

\newpage
## Interpolation
```{r new grid, eval=TRUE, echo=FALSE, warnings=FALSE, fig.align="center"}
# Set up grid, for prediction
grid <- rasterToPoints(aggregate(landDEM, fact=200)) #, fact=20))
grid <- grid[,1:2]
grid <- SpatialPointsDataFrame(grid,data.frame(elev=extract(landDEM,grid)))
grid$cont <- extract(ocean.dist, grid)
grid$hsun <- extract(h.w, grid)
grid$dist <- extract(atmo.dist.w, grid)

# Define RMSE function
RMSE <- function(x,y){
  tmp <- (x-y)^2
  sqrt(mean(tmp))
}
```

### Winter before 1970
```{r temp1970wUOK, eval=TRUE, echo=FALSE, warnings=FALSE, fig.align="center"}
# Extract continentality and elevation at measurement points from rasters
temp1970w$elev <- extract(landDEM, temp1970w)
temp1970w$elev[is.na(temp1970w$elev)] <- 0
temp1970w$cont <- extract(ocean.dist, temp1970w)
temp1970w$cont[is.na(temp1970w$cont)] <- 0
temp1970w$hsun <- extract(h.w, temp1970w)
temp1970w$hsun[is.na(temp1970w$hsun)] <- 0
temp1970w$dist <- extract(atmo.dist.w, temp1970w)
temp1970w$dist[is.na(temp1970w$dist)] <- 0

# Linear Regression
temp1970w.lm.test <- lm(meansum~elev+cont+hsun+dist, data=temp1970w@data)
summary(temp1970w.lm.test)

# Distribution of residuals
temp1970w@data[names(temp1970w.lm.test$residuals),"lmRes"]<-temp1970w.lm.test$residuals
temp1970w@data$lmRes[is.na(temp1970w@data$lmRes)] <- 0
temp1970w$lmResRel<-temp1970w$lmRes/temp1970w$meansum

# Plot results
bubble(temp1970w,"lmRes",main="Residual Values", na.rm=T)
bubble(temp1970w,"lmResRel",main="Relative Residual Values", na.rm=T)
# plot(variogram(lmRes~1,temp1970w,width=10,cutoff=200),main="Residual Variogram")

# Compute distance matrix
temp1970w.d <- as.matrix(dist(cbind(temp1970w@coords[,1], temp1970w@coords[,2])), method = "euclidean", alternative = "greater")

# Inverse distance matrix
temp1970w.d.inv <- 1 / temp1970w.d

# Setting diagonal to 0
diag(temp1970w.d.inv) <- 0

# Print Moran's I
temp.moran <- Moran.I(temp1970w$lmRes, temp1970w.d.inv)
paste("Observed autocorrelation: ",temp.moran$observed)
paste("P-value of H0 (residuals are randomly distributed): ",temp.moran$p.value)

# Universal Kriging
temp1970w.best.m.index <- 3 # Gaussian model
temp1970w.uok <- krige(meansum~elev+cont+hsun+dist,
                       temp1970w, grid,
                       model = temp1970w.var10.fits[[temp1970w.best.m.index]])

# Create grid
gridded(temp1970w.uok)<-TRUE
temp1970w.uok.pred<-raster(temp1970w.uok, layer=1, values=TRUE)
temp1970w.uok.var<-raster(temp1970w.uok, layer=2, values=TRUE)

# Extract and append to test data
# "meanWi_before1970" "meanSu_before1970" "meanWi_after1990"  "meanSu_after1990"
temp.test$meanWi_before1970 <- extract(temp1970w.uok.pred, temp.test)

# Calculate RMSE of map
temp.val$meanWi_before1970_pred <- extract(temp1970w.uok.pred, temp.val)
paste("Observed RMSE (5% validation data): ", round(RMSE(temp.val$meanWi_before1970, temp.val$meanWi_before1970_pred),2),"°C", sep="")

# Plot
p1 <- levelplot(temp1970w.uok.pred,
          main = "Prediction: Winter before 1970",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(temp1970w.uok.pred)), maxValue(abs(temp1970w.uok.pred)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p1)

p2 <- levelplot(temp1970w.uok.var,
          main = "Uncertainty: Winter before 1970",
          par.settings=YlOrRdTheme(), margin=F, 
          at=seq(0, maxValue(abs(temp1970w.uok.var)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p2)

# Save plots
pdf(file.path(figureFolder, "temp1970w_pred.pdf"), width=10.5, height=7.5)
print(p1)
invisible(dev.off())

pdf(file.path(figureFolder, "temp1970w_var.pdf"), width=10.5, height=7.5)
print(p2)
invisible(dev.off())

# Clear plots
remove(p1, p2)

# Save raster
writeRaster(temp1970w.uok.pred, file.path(resultsFolder, "temp1970w_pred"), format = "GTiff", overwrite=TRUE)
writeRaster(temp1970w.uok.var, file.path(resultsFolder, "temp1970w_var"), format = "GTiff", overwrite=TRUE)
```

\newpage
### Winter after 1990
```{r temp2010wUOK, eval=TRUE, echo=FALSE, warnings=FALSE, fig.align="center"}
# Extract continentality and elevation at measurement points from rasters
temp2010w$elev <- extract(landDEM, temp2010w)
temp2010w$elev[is.na(temp2010w$elev)] <- 0
temp2010w$cont <- extract(ocean.dist, temp2010w)
temp2010w$cont[is.na(temp2010w$cont)] <- 0
temp2010w$hsun <- extract(h.w, temp2010w)
temp2010w$hsun[is.na(temp2010w$hsun)] <- 0
temp2010w$dist <- extract(atmo.dist.w, temp2010w)
temp2010w$dist[is.na(temp2010w$dist)] <- 0

# Linear Regression
temp2010w.lm.test <- lm(meansum~elev+cont+hsun+dist, data=temp2010w@data)
summary(temp2010w.lm.test)

# Distribution of residuals
temp2010w@data[names(temp2010w.lm.test$residuals),"lmRes"]<-temp2010w.lm.test$residuals
temp2010w@data$lmRes[is.na(temp2010w@data$lmRes)] <- 0
temp2010w$lmResRel<-temp2010w$lmRes/temp2010w$meansum

# Remove outlier
temp2010w$lmResRel[which.max(temp2010w$lmResRel)] <- mean(temp2010w$lmResRel)
temp2010w$lmResRel[which.max(temp2010w$lmResRel)] <- mean(temp2010w$lmResRel)
temp2010w$lmResRel[which.max(temp2010w$lmResRel)] <- mean(temp2010w$lmResRel)
temp2010w$lmResRel[which.max(temp2010w$lmResRel)] <- mean(temp2010w$lmResRel)
temp2010w$lmResRel[which.max(temp2010w$lmResRel)] <- mean(temp2010w$lmResRel)

# Plot results
bubble(temp2010w,"lmRes",main="Residual Values", na.rm=T)
bubble(temp2010w,"lmResRel",main="Relative Residual Values", na.rm=T)

# Compute distance matrix
temp2010w.d <- as.matrix(dist(cbind(temp2010w@coords[,1], temp2010w@coords[,2])), method = "euclidean", alternative = "greater")

# Inverse distance matrix
temp2010w.d.inv <- 1 / temp2010w.d

# Setting diagonal to 0
diag(temp2010w.d.inv) <- 0

# Print Moran's I
temp.moran <- Moran.I(temp2010w$lmRes, temp2010w.d.inv)
paste("Observed autocorrelation: ",temp.moran$observed)
paste("P-value of H0 (residuals are randomly distributed): ",temp.moran$p.value)

# Universal Kriging
temp2010w.best.m.index <- 3 # Gaussian model
temp2010w.uok <- krige(meansum~elev+cont+hsun+dist,
                       temp2010w, grid,
                       model = temp2010w.var10.fits[[temp2010w.best.m.index]])

# Create grid
gridded(temp2010w.uok)<-TRUE
temp2010w.uok.pred<-raster(temp2010w.uok, layer=1, values=TRUE)
temp2010w.uok.var<-raster(temp2010w.uok, layer=2, values=TRUE)

# Extract and append to test data
# "meanWi_before1970" "meanSu_before1970" "meanWi_after1990"  "meanSu_after1990"
temp.test$meanWi_after1990 <- extract(temp2010w.uok.pred, temp.test)

# Calculate RMSE of map
temp.val$meanWi_after1990_pred <- extract(temp2010w.uok.pred, temp.val)
paste("Observed RMSE (5% validation data): ", round(RMSE(temp.val$meanWi_after1990, temp.val$meanWi_after1990_pred),2),"°C", sep="")

# Plot
p1 <- levelplot(temp2010w.uok.pred,
          main = "Prediction: Winter after 1990",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(temp2010w.uok.pred)), maxValue(abs(temp2010w.uok.pred)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p1)

p2 <- levelplot(temp2010w.uok.var,
          main = "Uncertainty: Winter after 1990",
          par.settings=YlOrRdTheme(), margin=F, 
          at=seq(0, maxValue(abs(temp2010w.uok.var)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p2)

# Save plots
pdf(file.path(figureFolder, "temp2010w_pred.pdf"), width=10.5, height=7.5)
print(p1)
invisible(dev.off())

pdf(file.path(figureFolder, "temp2010w_var.pdf"), width=10.5, height=7.5)
print(p2)
invisible(dev.off())

# Clear plots
remove(p1, p2)

# Save raster
writeRaster(temp2010w.uok.pred, file.path(resultsFolder, "temp2010w_pred"), format = "GTiff", overwrite=TRUE)
writeRaster(temp2010w.uok.var, file.path(resultsFolder, "temp2010w_var"), format = "GTiff", overwrite=TRUE)
```

\newpage
### Summer before 1970
```{r temp1970sUOK, eval=TRUE, echo=FALSE, warnings=FALSE, fig.align="center"}
# !!! Adjust grid for summer
grid$hsun <- extract(h.s, grid)
grid$dist <- extract(atmo.dist.s, grid)

# Extract continentality and elevation at measurement points from rasters
temp1970s$elev <- extract(landDEM, temp1970s)
temp1970s$elev[is.na(temp1970s$elev)] <- 0
temp1970s$cont <- extract(ocean.dist, temp1970s)
temp1970s$cont[is.na(temp1970s$cont)] <- 0
temp1970s$hsun <- extract(h.s, temp1970s)
temp1970s$hsun[is.na(temp1970s$hsun)] <- 0
temp1970s$dist <- extract(atmo.dist.s, temp1970s)
temp1970s$dist[is.na(temp1970s$dist)] <- 0

# Linear Regression
temp1970s.lm.test <- lm(meansum~elev+cont+hsun+dist, data=temp1970s@data)
summary(temp1970s.lm.test)

# Distribution of residuals
temp1970s@data[names(temp1970s.lm.test$residuals),"lmRes"]<-temp1970s.lm.test$residuals
temp1970s@data$lmRes[is.na(temp1970s@data$lmRes)] <- 0
temp1970s$lmResRel<-temp1970s$lmRes/temp1970s$meansum

# Plot results
bubble(temp1970s,"lmRes",main="Residual Values", na.rm=T)
bubble(temp1970s,"lmResRel",main="Relative Residual Values", na.rm=T)
# plot(variogram(lmRes~1,temp1970s,width=10,cutoff=200),main="Residual Variogram")

# Compute distance matrix
temp1970s.d <- as.matrix(dist(cbind(temp1970s@coords[,1], temp1970s@coords[,2])), method = "euclidean", alternative = "greater")

# Inverse distance matrix
temp1970s.d.inv <- 1 / temp1970s.d

# Setting diagonal to 0
diag(temp1970s.d.inv) <- 0

# Print Moran's I
temp.moran <- Moran.I(temp1970s$lmRes, temp1970s.d.inv)
paste("Observed autocorrelation: ",temp.moran$observed)
paste("P-value of H0 (residuals are randomly distributed): ",temp.moran$p.value)

# Universal Kriging
temp1970s.best.m.index <- 3 # Gaussian model
temp1970s.uok <- krige(meansum~elev+cont+hsun+dist,
                       temp1970s, grid,
                       model = temp1970s.var10.fits[[temp1970s.best.m.index]])

# Create grid
gridded(temp1970s.uok)<-TRUE
temp1970s.uok.pred<-raster(temp1970s.uok, layer=1, values=TRUE)
temp1970s.uok.var<-raster(temp1970s.uok, layer=2, values=TRUE)

# Extract and append to test data
# "meanWi_before1970" "meanSu_before1970" "meanWi_after1990"  "meanSu_after1990"
temp.test$meanSu_before1970 <- extract(temp1970s.uok.pred, temp.test)

# Calculate RMSE of map
temp.val$meanSu_before1970_pred <- extract(temp1970s.uok.pred, temp.val)
paste("Observed RMSE (5% validation data): ", round(RMSE(temp.val$meanSu_before1970, temp.val$meanSu_before1970_pred),2),"°C", sep="")

# Plot
p1 <- levelplot(temp1970s.uok.pred,
          main = "Prediction: Summer before 1970",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(temp1970s.uok.pred)), maxValue(abs(temp1970s.uok.pred)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p1)

p2 <- levelplot(temp1970s.uok.var,
          main = "Uncertainty: Summer before 1970",
          par.settings=YlOrRdTheme(), margin=F, 
          at=seq(0, maxValue(abs(temp1970s.uok.var)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p2)

# Save plots
pdf(file.path(figureFolder, "temp1970s_pred.pdf"), width=10.5, height=7.5)
print(p1)
invisible(dev.off())

pdf(file.path(figureFolder, "temp1970s_var.pdf"), width=10.5, height=7.5)
print(p2)
invisible(dev.off())

# Clear plots
remove(p1, p2)

# Save raster
writeRaster(temp1970s.uok.pred, file.path(resultsFolder, "temp1970s_pred"), format = "GTiff", overwrite=TRUE)
writeRaster(temp1970s.uok.var, file.path(resultsFolder, "temp1970s_var"), format = "GTiff", overwrite=TRUE)
```

\newpage
### Summer after 1990
```{r temp2010sUOK, eval=TRUE, echo=FALSE, warnings=FALSE, fig.align="center"}
# Extract continentality and elevation at measurement points from rasters
temp2010s$elev <- extract(landDEM, temp2010s)
temp2010s$elev[is.na(temp2010s$elev)] <- 0
temp2010s$cont <- extract(ocean.dist, temp2010s)
temp2010s$cont[is.na(temp2010s$cont)] <- 0
temp2010s$hsun <- extract(h.s, temp2010s)
temp2010s$hsun[is.na(temp2010s$hsun)] <- 0
temp2010s$dist <- extract(atmo.dist.s, temp2010s)
temp2010s$dist[is.na(temp2010s$dist)] <- 0

# Linear Regression
temp2010s.lm.test <- lm(meansum~elev+cont+hsun+dist, data=temp2010s@data)
summary(temp2010s.lm.test)

# Distribution of residuals
temp2010s@data[names(temp2010s.lm.test$residuals),"lmRes"]<-temp2010s.lm.test$residuals
temp2010s@data$lmRes[is.na(temp2010s@data$lmRes)] <- 0
temp2010s$lmResRel<-temp2010s$lmRes/temp2010s$meansum

# Plot results
bubble(temp2010s,"lmRes",main="Residual Values", na.rm=T)
bubble(temp2010s,"lmResRel",main="Relative Residual Values", na.rm=T)
# plot(variogram(lmRes~1,temp2010s,width=10,cutoff=200),main="Residual Variogram")

# Compute distance matrix
temp2010s.d <- as.matrix(dist(cbind(temp2010s@coords[,1], temp2010s@coords[,2])), method = "euclidean", alternative = "greater")

# Inverse distance matrix
temp2010s.d.inv <- 1 / temp2010s.d

# Setting diagonal to 0
diag(temp2010s.d.inv) <- 0

# Print Moran's I
temp.moran <- Moran.I(temp2010s$lmRes, temp2010s.d.inv)
paste("Observed autocorrelation: ",temp.moran$observed)
paste("P-value of H0 (residuals are randomly distributed): ",temp.moran$p.value)

# Universal Kriging
temp2010s.best.m.index <- 3 # Gaussian model
temp2010s.uok <- krige(meansum~elev+cont+hsun+dist,
                       temp2010s, grid,
                       model = temp2010s.var10.fits[[temp2010s.best.m.index]])

# Create grid
gridded(temp2010s.uok)<-TRUE
temp2010s.uok.pred<-raster(temp2010s.uok, layer=1, values=TRUE)
temp2010s.uok.var<-raster(temp2010s.uok, layer=2, values=TRUE)

# Extract and append to test data
# "meanWi_before1970" "meanSu_before1970" "meanWi_after1990"  "meanSu_after1990"
temp.test$meanSu_after1990 <- extract(temp2010s.uok.pred, temp.test)

# Calculate RMSE of map
temp.val$meanSu_after1990_pred <- extract(temp2010s.uok.pred, temp.val)
paste("Observed RMSE (5% validation data): ", round(RMSE(temp.val$meanSu_after1990, temp.val$meanSu_after1990_pred),2),"°C", sep="")

# Plot
p1 <- levelplot(temp2010s.uok.pred,
          main = "Prediction: Summer after 1990",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(temp2010s.uok.pred)), maxValue(abs(temp2010s.uok.pred)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p1)

p2 <- levelplot(temp2010s.uok.var,
          main = "Uncertainty: Summer after 1990",
          par.settings=YlOrRdTheme(), margin=F, 
          at=seq(0, maxValue(abs(temp2010s.uok.var)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p2)

# Save plots
pdf(file.path(figureFolder, "temp2010s_pred.pdf"), width=10.5, height=7.5)
print(p1)
invisible(dev.off())

pdf(file.path(figureFolder, "temp2010s_var.pdf"), width=10.5, height=7.5)
print(p2)
invisible(dev.off())

# Clear plots
remove(p1, p2)

# Save raster
writeRaster(temp2010s.uok.pred, file.path(resultsFolder, "temp2010s_pred"), format = "GTiff", overwrite=TRUE)
writeRaster(temp2010s.uok.var, file.path(resultsFolder, "temp2010s_var"), format = "GTiff", overwrite=TRUE)

# Save temperature_test_pred.csv
write.csv(temp.test, file = file.path(resultsFolder, "temperature_test_pred"), row.names = FALSE)
```

\newpage
# Difference images
## Winter
```{r winterdiff, eval=TRUE, echo=FALSE, warnings=FALSE, fig.align="center"}
diff.w <- temp2010w.uok.pred-temp1970w.uok.pred

# Plot
p <- levelplot(diff.w,
          main = "Temperature change: Winter",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(diff.w)), maxValue(abs(diff.w)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p)

# Save plots
pdf(file.path(figureFolder, "diff_winter.pdf"), width=10.5, height=7.5)
print(p)
invisible(dev.off())

# Clear plots
remove(p)

# Save raster
writeRaster(diff.w, file.path(resultsFolder, "diff_winter"), format = "GTiff", overwrite=TRUE)
```

## Summer
```{r summerdiff, eval=TRUE, echo=FALSE, warnings=FALSE, fig.align="center"}
diff.s <- temp2010s.uok.pred-temp1970s.uok.pred

# Plot
p <- levelplot(diff.s,
          main = "Temperature change: Summer",
          par.settings=BuRdTheme(), margin=F, 
          at=seq(-maxValue(abs(diff.s)), maxValue(abs(diff.s)), len=100)) + 
  layer(sp.polygons(oceans, fill='transparent'))
print(p)

# Save plots
pdf(file.path(figureFolder, "diff_summer.pdf"), width=10.5, height=7.5)
print(p)
invisible(dev.off())

# Clear plots
remove(p)

# Save raster
writeRaster(diff.s, file.path(resultsFolder, "diff_summer"), format = "GTiff", overwrite=TRUE)
```
