---
title: 'Challenge II: Spatial continuity and weather prediction'
author:
- "Olivier Niklaus, Geography UZH"
- "Merlin Unterfinger, Geography UZH"
date: 25 6 2017
output:
  rmarkdown::pdf_document:
    fig_caption: yes        
    includes:  
      in_header: preamble-latex.tex
    number_sections: yes
    toc: yes
    toc_depth: 3
#    citation_package: natbib
  html_document: default
#biblio-style: "apalike"
#link-citations: true
#csl: apa.csl
#bibliography: References.bib
#abstract: "blablabalbalbalbalbalbalbalbalbalaalbalbla"
filter: pandoc-eqnos
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())             # Clean the environment
options(scipen=6)         # Display digits, not the scientific version
options(digits.secs=6)    # use milliseconds in Date/Time data types
options(warning=FALSE)    # Don't show warnings
par(mfrow=c(1,1))         # reset plot placement to normal 1 by 1

dataFolder    <- file.path(getwd(), "data")            # Data folder
figureFolder  <- file.path(getwd(), "figures")         # Figure Folder
```

\newpage

# Libraries
The following libraries are needed, to execute the code:

* **ggplot2** -- Plots -- [@ggplot2]
* **plyr** -- Splitting, applying and combining data -- [@plyr]
* **rgdal** -- Open Shapefiles -- [@rgdal]
* **sp** -- Spatial datatypes -- [@sp2; @sp1]
* **gstat** -- Spatial statistics --[@gstat]
* **FNN** -- Spatial statistics --[@FNN]


```{r Libraries, message=FALSE, echo=FALSE}
#library(e1071)            # Support Vector Machine library
library(sp)               # Spatial data types
#library(rgeos)            # Geoprocessing
library(rgdal)            # Open Shapefiles
library(ggplot2)          # Plots
library(gdistance)        # Least cost path - Walking time
#library(rasterVis)        # Plots of rasters
library(plyr)
#spatial statistics library
#install.packages("gstat")
library(gstat)
#install.packages("FNN")
library(FNN)
#install.packages("spdep")
library(spdep)
library(gridExtra)

```

# Data
```{r Geodata, message=FALSE, warning=FALSE , echo=FALSE, fig.height=3, fig.align="center"}
# Define CRS
WGS84 <- CRS("+init=epsg:4326")

# Read CSV file for temperature
temp <- read.csv(file.path(dataFolder, "temperature.csv"))
temp <- temp[complete.cases(temp), ]
coordinates(temp)<-~long+lat
proj4string(temp)<-WGS84

# Read test file
temp.test <- read.csv(file.path(dataFolder, "temperature_test.csv"))
coordinates(temp.test) <- ~long+lat
proj4string(temp.test) <- WGS84

plot(temp)
points(temp.test, col="red")
knitr::kable(head(round(as.data.frame(head(temp)),2), caption = "Structure of the temperature data set.", row.names = FALSE))

# Seperate data according to timeFrame and season
temp1970w <-SpatialPointsDataFrame(coords = temp@coords, data = temp@data[,1:2])
temp1970s <-SpatialPointsDataFrame(coords = temp@coords, data = temp@data[,c(1,3)])
temp2010w <-SpatialPointsDataFrame(coords = temp@coords, data = temp@data[,c(1,4)])
temp2010s <-SpatialPointsDataFrame(coords = temp@coords, data = temp@data[,c(1,5)])

col <- c("id", "meansum")
colnames(temp1970s@data) <- col
colnames(temp2010s@data) <- col
colnames(temp1970w@data) <- col
colnames(temp2010w@data) <- col

# Open digital elevation model
DEM <-readGDAL(file.path(dataFolder, "globalDHM.tif"), silent = T)
plot(DEM)
```

\newpage
# Spatial continuity
## First impressions of spatial variation
### Winter before 1970
```{r temp1970w, eval=TRUE, echo=FALSE, fig.height=3, fig.align="center", fig.cap="Mean temperature Winter from 1950 to 1970"}

#plotting the coefficients of variation for all varibles
spplot(temp1970w, "meansum", colorkey=T)
```

### Summer before 1970
```{r temp1970s, eval=TRUE, echo=FALSE, fig.height=3, fig.align="center", fig.cap="Mean temperature Winter from 1950 to 1970"}

#plotting the coefficients of variation for all varibles
spplot(temp1970s,"meansum", colorkey=T)
```
### Winter after 1990
```{r temp2010w, eval=TRUE, echo=FALSE, fig.height=3, fig.align="center", fig.cap="Mean temperature Winter from 1990 to 2010"}

#plotting the coefficients of variation for all varibles
spplot(temp2010w,"meansum", colorkey=T)
```
### Summer after 1990
```{r temp2010s, eval=TRUE, echo=FALSE, fig.height=3, fig.align="center", fig.cap="Mean temperature Winter from 1990 to 2010"}

#plotting the coefficients of variation for all varibles
spplot(temp2010s,"meansum", colorkey=T)
```
## H - Scatterplots and autocovariance plots
### Summer before 1970
```{r scatter, echo=FALSE, fig.align="center", fig.height=4}

temp.dist<-spDists(temp1970s,longlat = FALSE)
temp.index <-knn.index(temp.dist, k=20, algorithm=c("kd_tree"))

temp.h<-data.frame(sum=temp1970s$meansum ,sumNN1=temp1970s$meansum[temp.index[,1]],sumNN5=temp1970s$meansum[temp.index[,5]],sumNN10=temp1970s$meansum[temp.index[,10]],sumNN20=temp1970s$meansum[temp.index[,20]])

par(mfrow=c(2,2))
plot(temp.h$sum,temp.h$sumNN1,main="h = 1")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN5,main="h = 5")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN10,main="h = 10")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN20,main="h = 20")
abline(1,1, col = "blue")

##Autocovariance plots
temp.mean <- mean(temp.h$sum, na.rm = TRUE)
temp.sd<-sd(temp.h$sum, na.rm = TRUE)

temp.acov.nn1<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN1-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn5<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN5-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn10<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN10-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn20<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN20-temp.mean), na.rm = TRUE)/nrow(temp.h)

temp.acorr.nn1<-temp.acov.nn1/temp.sd^2
temp.acorr.nn5<-temp.acov.nn5/temp.sd^2
temp.acorr.nn10<-temp.acov.nn10/temp.sd^2
temp.acorr.nn20<-temp.acov.nn20/temp.sd^2

par(mfrow=c(1,2))
plot(c(1,5,10,20),c(temp.acov.nn1,temp.acov.nn5,temp.acov.nn10,temp.acov.nn20),type="l",main="auto-covariance")
plot(c(1,5,10,20),c(temp.acorr.nn1,temp.acorr.nn5,temp.acorr.nn10,temp.acorr.nn20),type="l",main="auto-correlation")

```
### Winter before 1970
```{r scatter2, echo=FALSE, fig.align="center", fig.height=4}

temp.dist<-spDists(temp1970w,longlat = FALSE)
temp.index <-knn.index(temp.dist, k=20, algorithm=c("kd_tree"))

temp.h<-data.frame(sum=temp1970w$meansum ,sumNN1=temp1970w$meansum[temp.index[,1]],sumNN5=temp1970w$meansum[temp.index[,5]],sumNN10=temp1970w$meansum[temp.index[,10]],sumNN20=temp1970w$meansum[temp.index[,20]])

par(mfrow=c(2,2))
plot(temp.h$sum,temp.h$sumNN1,main="h = 1")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN5,main="h = 5")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN10,main="h = 10")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN20,main="h = 20")
abline(1,1, col = "blue")

##Autocoavariance plots
temp.mean<-mean(temp.h$sum, na.rm = TRUE)
temp.sd<-sd(temp.h$sum, na.rm = TRUE)

temp.acov.nn1<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN1-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn5<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN5-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn10<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN10-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn20<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN20-temp.mean), na.rm = TRUE)/nrow(temp.h)

temp.acorr.nn1<-temp.acov.nn1/temp.sd^2
temp.acorr.nn5<-temp.acov.nn5/temp.sd^2
temp.acorr.nn10<-temp.acov.nn10/temp.sd^2
temp.acorr.nn20<-temp.acov.nn20/temp.sd^2

par(mfrow=c(1,2))
plot(c(1,5,10,20),c(temp.acov.nn1,temp.acov.nn5,temp.acov.nn10,temp.acov.nn20),type="l",main="auto-covariance")
plot(c(1,5,10,20),c(temp.acorr.nn1,temp.acorr.nn5,temp.acorr.nn10,temp.acorr.nn20),type="l",main="auto-correlation")

```
### Winter after 1990
```{r scatter3, echo=FALSE, fig.align="center", fig.height=4}

temp.dist<-spDists(temp2010w,longlat = FALSE)
temp.index <-knn.index(temp.dist, k=20, algorithm=c("kd_tree"))

temp.h<-data.frame(sum=temp2010w$meansum ,sumNN1=temp2010w$meansum[temp.index[,1]],sumNN5=temp2010w$meansum[temp.index[,5]],sumNN10=temp2010w$meansum[temp.index[,10]],sumNN20=temp2010w$meansum[temp.index[,20]])

par(mfrow=c(2,2))
plot(temp.h$sum,temp.h$sumNN1,main="h = 1")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN5,main="h = 5")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN10,main="h = 10")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN20,main="h = 20")
abline(1,1, col = "blue")

##Autocoavariance plots
temp.mean<-mean(temp.h$sum, na.rm = TRUE)
temp.sd<-sd(temp.h$sum, na.rm = TRUE)

temp.acov.nn1<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN1-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn5<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN5-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn10<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN10-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn20<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN20-temp.mean), na.rm = TRUE)/nrow(temp.h)

temp.acorr.nn1<-temp.acov.nn1/temp.sd^2
temp.acorr.nn5<-temp.acov.nn5/temp.sd^2
temp.acorr.nn10<-temp.acov.nn10/temp.sd^2
temp.acorr.nn20<-temp.acov.nn20/temp.sd^2

par(mfrow=c(1,2))
plot(c(1,5,10,20),c(temp.acov.nn1,temp.acov.nn5,temp.acov.nn10,temp.acov.nn20),type="l",main="auto-covariance")
plot(c(1,5,10,20),c(temp.acorr.nn1,temp.acorr.nn5,temp.acorr.nn10,temp.acorr.nn20),type="l",main="auto-correlation")
```
### Summer after 1990
```{r scatter4, echo=FALSE, fig.align="center", fig.height=4}

temp.dist<-spDists(temp2010s,longlat = FALSE)
temp.index <-knn.index(temp.dist, k=20, algorithm=c("kd_tree"))

temp.h<-data.frame(sum=temp2010s$meansum ,sumNN1=temp2010s$meansum[temp.index[,1]],sumNN5=temp2010s$meansum[temp.index[,5]],sumNN10=temp2010s$meansum[temp.index[,10]],sumNN20=temp2010s$meansum[temp.index[,20]])

par(mfrow=c(2,2))
plot(temp.h$sum,temp.h$sumNN1,main="h = 1")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN5,main="h = 5")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN10,main="h = 10")
abline(1,1, col = "blue")
plot(temp.h$sum,temp.h$sumNN20,main="h = 20")
abline(1,1, col = "blue")

##Autocoavariance plots
temp.mean<-mean(temp.h$sum, na.rm = TRUE)
temp.sd<-sd(temp.h$sum, na.rm = TRUE)

temp.acov.nn1<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN1-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn5<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN5-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn10<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN10-temp.mean), na.rm = TRUE)/nrow(temp.h)
temp.acov.nn20<-sum((temp.h$sum-temp.mean)*(temp.h$sumNN20-temp.mean), na.rm = TRUE)/nrow(temp.h)

temp.acorr.nn1<-temp.acov.nn1/temp.sd^2
temp.acorr.nn5<-temp.acov.nn5/temp.sd^2
temp.acorr.nn10<-temp.acov.nn10/temp.sd^2
temp.acorr.nn20<-temp.acov.nn20/temp.sd^2

par(mfrow=c(1,2))
plot(c(1,5,10,20),c(temp.acov.nn1,temp.acov.nn5,temp.acov.nn10,temp.acov.nn20),type="l",main="auto-covariance")
plot(c(1,5,10,20),c(temp.acorr.nn1,temp.acorr.nn5,temp.acorr.nn10,temp.acorr.nn20),type="l",main="auto-correlation")

```

## Empirical Variogram
### Winter before 1970
```{r temp1970wvar, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp1970w.var1<-variogram(meansum~1,temp1970w,cutoff=1000,width=1)
temp1970w.var5<-variogram(meansum~1,temp1970w,cutoff=150,width=5)
temp1970w.var10<-variogram(meansum~1,temp1970w,cutoff=150,width=10)
temp1970w.var15<-variogram(meansum~1,temp1970w,cutoff=150,width=15)

grid.arrange(plot(temp1970w.var1, main="Bin width 1km"), 
             plot(temp1970w.var5, main="Bin width 5km"), ncol = 2)

grid.arrange(plot(temp1970w.var10, main="Bin width 10km"), 
             plot(temp1970w.var15, main="Bin width 15km"), ncol = 2)

bws<-data.frame(np=temp1970w.var1$np,bw=rep("1km",length(temp1970w.var1$np)))
bws<-rbind(bws,data.frame(np=temp1970w.var5$np,bw=rep("10km",length(temp1970w.var5$np))))
bws<-rbind(bws,data.frame(np=temp1970w.var10$np,bw=rep("20km",length(temp1970w.var10$np))))
bws<-rbind(bws,data.frame(np=temp1970w.var15$np,bw=rep("30km",length(temp1970w.var15$np))))

#plotting the numbers of points per bin and bandwidth as a boxplot
ggplot(bws, aes(x=bw, y=np, color=bw)) +
  geom_boxplot()
```

\newpage
### Summer before 1970
```{r temp1970svar, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp1970s.var1<-variogram(meansum~1,temp1970s,cutoff=1000,width=1)
temp1970s.var5<-variogram(meansum~1,temp1970s,cutoff=150,width=5)
temp1970s.var10<-variogram(meansum~1,temp1970s,cutoff=150,width=10)
temp1970s.var15<-variogram(meansum~1,temp1970s,cutoff=150,width=15)

grid.arrange(plot(temp1970s.var1, main="Bin width 1km"), 
             plot(temp1970s.var5, main="Bin width 5km"), ncol = 2)

grid.arrange(plot(temp1970s.var10, main="Bin width 10km"), 
             plot(temp1970s.var15, main="Bin width 15km"), ncol = 2)

bws<-data.frame(np=temp1970s.var1$np,bw=rep("1km",length(temp1970s.var1$np)))
bws<-rbind(bws,data.frame(np=temp1970s.var5$np,bw=rep("10km",length(temp1970s.var5$np))))
bws<-rbind(bws,data.frame(np=temp1970s.var10$np,bw=rep("20km",length(temp1970s.var10$np))))
bws<-rbind(bws,data.frame(np=temp1970s.var15$np,bw=rep("30km",length(temp1970s.var15$np))))

#plotting the numbers of points per bin and bandwidth as a boxplot
ggplot(bws, aes(x=bw, y=np, color=bw)) +
  geom_boxplot()
```

\newpage
### Winter before 2010
```{r temp2010wvar, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp2010w.var1<-variogram(meansum~1,temp2010w,cutoff=1000,width=1)
temp2010w.var5<-variogram(meansum~1,temp2010w,cutoff=150,width=5)
temp2010w.var10<-variogram(meansum~1,temp2010w,cutoff=150,width=10)
temp2010w.var15<-variogram(meansum~1,temp2010w,cutoff=150,width=15)

grid.arrange(plot(temp2010w.var1, main="Bin width 1km"), 
             plot(temp2010w.var5, main="Bin width 5km"), ncol = 2)

grid.arrange(plot(temp2010w.var10, main="Bin width 10km"), 
             plot(temp2010w.var15, main="Bin width 15km"), ncol = 2)

bws<-data.frame(np=temp2010w.var1$np,bw=rep("1km",length(temp2010w.var1$np)))
bws<-rbind(bws,data.frame(np=temp2010w.var5$np,bw=rep("10km",length(temp2010w.var5$np))))
bws<-rbind(bws,data.frame(np=temp2010w.var10$np,bw=rep("20km",length(temp2010w.var10$np))))
bws<-rbind(bws,data.frame(np=temp2010w.var15$np,bw=rep("30km",length(temp2010w.var15$np))))

#plotting the numbers of points per bin and bandwidth as a boxplot
ggplot(bws, aes(x=bw, y=np, color=bw)) +
  geom_boxplot()
```

\newpage
### Summer before 2010
```{r temp2010svar, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp2010s.var1<-variogram(meansum~1,temp2010s,cutoff=1000,width=1)
temp2010s.var5<-variogram(meansum~1,temp2010s,cutoff=150,width=5)
temp2010s.var10<-variogram(meansum~1,temp2010s,cutoff=150,width=10)
temp2010s.var15<-variogram(meansum~1,temp2010s,cutoff=150,width=15)

grid.arrange(plot(temp2010s.var1, main="Bin width 1km"), 
             plot(temp2010s.var5, main="Bin width 5km"), ncol = 2)

grid.arrange(plot(temp2010s.var10, main="Bin width 10km"), 
             plot(temp2010s.var15, main="Bin width 15km"), ncol = 2)

bws<-data.frame(np=temp2010s.var1$np,bw=rep("1km",length(temp2010s.var1$np)))
bws<-rbind(bws,data.frame(np=temp2010s.var5$np,bw=rep("10km",length(temp2010s.var5$np))))
bws<-rbind(bws,data.frame(np=temp2010s.var10$np,bw=rep("20km",length(temp2010s.var10$np))))
bws<-rbind(bws,data.frame(np=temp2010s.var15$np,bw=rep("30km",length(temp2010s.var15$np))))

#plotting the numbers of points per bin and bandwidth as a boxplot
ggplot(bws, aes(x=bw, y=np, color=bw)) +
  geom_boxplot()
```

## Fitted Semivariogram
### Winter before 1970
```{r temp1970wfit, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp1970w.var10<-variogram(meansum~1,temp1970w,cutoff=150,width=10)

models <- c("Exp", "Sph", "Gau", "Mat")
temp1970w.var10.fits <- lapply(X=1:4, FUN=function(x) fit.variogram(temp1970w.var10, model=vgm(models[x])))

grid.arrange(plot(temp1970w.var10, model=temp1970w.var10.fits[[1]], main = "Exponential: 10km"), 
             plot(temp1970w.var10, model=temp1970w.var10.fits[[2]], main = "Spherical: 10km"), ncol = 2)
grid.arrange(plot(temp1970w.var10, model=temp1970w.var10.fits[[3]], main = "Gaussian: 10km"), 
             plot(temp1970w.var10, model=temp1970w.var10.fits[[4]], main = "Mat: 10km"), ncol = 2)
```

\newpage
### Summer before 1970
```{r temp1970sfit, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp1970s.var10<-variogram(meansum~1,temp1970s,cutoff=150,width=10)

models <- c("Exp", "Sph", "Gau", "Mat")
temp1970s.var10.fits <- lapply(X=1:4, FUN=function(x) fit.variogram(temp1970s.var10, model=vgm(models[x])))

grid.arrange(plot(temp1970s.var10, model=temp1970s.var10.fits[[1]], main = "Exponential: 10km"), 
             plot(temp1970s.var10, model=temp1970s.var10.fits[[2]], main = "Spherical: 10km"), ncol = 2)
grid.arrange(plot(temp1970s.var10, model=temp1970s.var10.fits[[3]], main = "Gaussian: 10km"), 
             plot(temp1970s.var10, model=temp1970s.var10.fits[[4]], main = "Mat: 10km"), ncol = 2)
```

\newpage
### Winter after 1990
```{r temp2010wfit, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp2010w.var10<-variogram(meansum~1,temp2010w,cutoff=150,width=10)

models <- c("Exp", "Sph", "Gau", "Mat")
temp2010w.var10.fits <- lapply(X=1:4, FUN=function(x) fit.variogram(temp2010w.var10, model=vgm(models[x])))

grid.arrange(plot(temp2010w.var10, model=temp2010w.var10.fits[[1]], main = "Exponential: 10km"), 
             plot(temp2010w.var10, model=temp2010w.var10.fits[[2]], main = "Spherical: 10km"), ncol = 2)
grid.arrange(plot(temp2010w.var10, model=temp2010w.var10.fits[[3]], main = "Gaussian: 10km"), 
             plot(temp2010w.var10, model=temp2010w.var10.fits[[4]], main = "Mat: 10km"), ncol = 2)
```

\newpage
### Summer after 1990
```{r temp2010sfit, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
temp2010s.var10<-variogram(meansum~1,temp2010s,cutoff=150,width=10)

models <- c("Exp", "Sph", "Gau", "Mat")
temp2010s.var10.fits <- lapply(X=1:4, FUN=function(x) fit.variogram(temp2010s.var10, model=vgm(models[x])))

grid.arrange(plot(temp2010s.var10, model=temp2010s.var10.fits[[1]], main = "Exponential: 10km"), 
             plot(temp2010s.var10, model=temp2010s.var10.fits[[2]], main = "Spherical: 10km"), ncol = 2)
grid.arrange(plot(temp2010s.var10, model=temp2010s.var10.fits[[3]], main = "Gaussian: 10km"), 
             plot(temp2010s.var10, model=temp2010s.var10.fits[[4]], main = "Mat: 10km"), ncol = 2)
```

\newpage
## Ordinary krigin
### Winter before 1970
```{r temp1970OK, eval=TRUE, echo=FALSE, fig.height=2.5, fig.align="center"}
# Use Gaussian model, with 10km bin width
temp1970w.var10.fits[[3]]

```
### Summer before 1970
### Winter after 1990
### Summer after 1990

## Ordinary Krigin
### Winter before 1970
### Summer before 1970
### Winter after 1990
### Summer after 1990

